\pkg{digiRhythm}: an \proglang{R} package for rhythmic behavior assessment in livestock

### 2023-05-07 jstatsoft/article_comr.Rnw
Marie_little changs

%% - \Abstract{} almost as usual # comr: changed abstract throughout.
\Abstract{
The welfare of farm animals is a growing concern to the public, yet indicators that allow to reliably quantify the welfare of these animals are scarce. One method that has been proposed to serve this purpose is measuring the organisms rhythmicity from activities acquired with sensors (e.g., accelerometers). The Degree of Functional Coupling (DFC) is one of these rhythmicity indicators. Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks. In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from activity data. Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of a diurnality index are also integrated in this package. The package is customized for integration in the research workflow, from raw data to creating publication-ready content and therefore allows the simple application of these sophisticated methods.
}

@@

The relationship between rhythmicity and frequencies would now be clear for the reader. If the significant frequencies are only the harmonic ones (i.e., ssh = sumsig), then the DFC = 1, and therefore the animal exhibits a highly rhythmic behavior. On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior., which in turn would indicate a poor physiological state.

>>>>

%% - \Abstract{} almost as usual # comr: changed abstract throughout.
\Abstract{
The welfare of farm animals is a growing concern to the public, yet indicators that allow to reliably quantify the welfare of these animals are scarce. One method that has been proposed to serve this purpose is measuring the organisms rhythmicity from activities acquired with sensors (e.g., accelerometers). The Degree of Functional Coupling (DFC) is one of these rhythmicity indicators. Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks. Additionally, categorizing the significant peaks in frequencies that are harmonically synchronised with the periodicity of the environment. In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from activity data. Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of a diurnality index are also integrated in this package. The package is customized for integration in the research workflow, from raw data to creating publication-ready content and therefore allows the simple application of these sophisticated methods.
}

@@

The relationship between rhythmicity and frequencies would now be clear for the reader. If the significant frequencies are only the harmonic ones (i.e., ssh = sumsig), then the DFC = 1, and therefore the animal exhibits a highly rhythmic behavior. On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior, which in turn could be initiated by a poor physiological state or other physiological events.
